Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ {\smaller \textbf{EIP-150 revision (\YellowPaperVersionNumber{})}}

### 2017-09-30 Paper.tex
Changed subscript to \mathrm e.g. for gas costs (using Word with Find and replace is much faster), added a citation for Mist.

\subsection{Execution}
We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{txdatazero} & \text{if} \quad i = 0 \\ G_{txdatanonzero} & \text{otherwise} \end{cases} \\
{} & + \begin{cases} G_\text{txcreate} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
{} & + G_{transaction}
\end{align}

where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call. $G_\text{txcreate}$ is added if the transaction is contract-creating, but not if a result of EVM-code. $G$ is fully defined in Appendix \ref{app:fees}.

@@

If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
\begin{equation}
c \equiv G_{codedeposit} \times |\mathbf{o}|
\end{equation}

@@

\subsection{Subtleties}
Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code. Thus any message call received by it during this time causes no code to be executed. If the initialisation execution ends with a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed. For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.

@@

The normal halting function $H$ is defined:
\begin{equation} \label{eq:Normal Halting}
H(\boldsymbol{\mu}, I) \equiv \begin{cases}
H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
\varnothing & \text{otherwise}
\end{cases}
\end{equation}

@@

\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol.

@@

\begin{tabularx}{\textwidth}{lrX}
\toprule
Name & Value & Description* \\
\midrule
$G_{zero}$ & 0 & Nothing is paid for operations of the set {\small $W_{zero}$}. \\
$G_{base}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{base}$}. \\
$G_{verylow}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{verylow}$}. \\
$G_{low}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{low}$}. \\
$G_{mid}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{mid}$}. \\
$G_{high}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{high}$}. \\
$G_{extcode}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{extcode}$}. \\
$G_{balance}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
$G_{sload}$ & 200 & This is paid for an {\small SLOAD} operation. \\
$G_{jumpdest}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
$G_{sset}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{sreset}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
$R_{sclear}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
$R_{selfdestruct}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
$G_{selfdestruct}$ & 5000 & This is the amount of gas to pay for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation. \\
$G_{create}$ & 32000 & This is paid for a {\small CREATE} operation. \\
$G_{codedeposit}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
$G_{call}$ & 700 & This is paid for a {\small CALL} operation. \\
$G_{callvalue}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
$G_{callstipend}$ & 2300 & This is a stipend for the called contract subtracted from $G_{callvalue}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
$G_{newaccount}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT} operation which creates an account. \\
$G_{exp}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
$G_{expbyte}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
$G_{memory}$ & 3 & This is paid for every additional word when expanding memory. \\
$G_\text{txcreate}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
$G_{txdatazero}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
$G_{txdatanonzero}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
$G_{transaction}$ & 21000 & This is paid for every transaction. \\
$G_{log}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
$G_{logdata}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
$G_{logtopic}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
$G_{sha3}$ & 30 & This is paid for each {\small SHA3} operation. \\
$G_{sha3word}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{copy}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
$G_{blockhash}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\

@@

\begin{equation}
C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathbf{i})-C_{mem}(\boldsymbol{\mu}_\mathbf{i}) + \begin{cases}
C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{exp} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
G_{exp} + G_{expbyte}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
G_{verylow} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
G_{extcode} + G_{copy}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{logtopic} & \text{if} \quad w = \text{\small LOG1} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{logtopic} & \text{if} \quad w = \text{\small LOG2} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{logtopic} & \text{if} \quad w = \text{\small LOG3} \\
G_{log}+G_{logdata}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{logtopic} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{SELFDESTRUCT}{\small SELFDESTRUCT}} \\
G_{create} & \text{if} \quad w = \text{\small CREATE}\\
G_{sha3}+G_{sha3word} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
G_{jumpdest} & \text{if} \quad w = \text{\small JUMPDEST}\\
G_{sload} & \text{if} \quad w = \text{\small SLOAD}\\
G_{zero} & \text{if} \quad w \in W_{zero}\\
G_{base} & \text{if} \quad w \in W_{base}\\
G_{verylow} & \text{if} \quad w \in W_{verylow}\\
G_{low} & \text{if} \quad w \in W_{low}\\
G_{mid} & \text{if} \quad w \in W_{mid}\\
G_{high} & \text{if} \quad w \in W_{high}\\
G_{extcode} & \text{if} \quad w \in W_{extcode}\\
G_{balance} & \text{if} \quad w = \text{\small BALANCE}\\
G_{blockhash} & \text{if} \quad w = \text{\small BLOCKHASH}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

where:
\begin{equation}
C_{mem}(a) \equiv G_{memory} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
\end{equation}

@@

$W_{zero}$ = \{{\small STOP}, {\small RETURN}\}

$W_{base}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}

$W_{verylow}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}

$W_{low}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}

$W_{mid}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}

$W_{high}$ = \{{\small JUMPI}\}

$W_{extcode}$ = \{{\small EXTCODESIZE}\}

Note the memory cost component, given as the product of $G_{memory}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write. Such accesses must be for non-zero number of bytes.

@@

G_{sset} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
G_{sreset} & \text{otherwise}

@@

R_{sclear} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\

@@

C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{callstipend} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\

@@

&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{call} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\

@@

G_{callvalue} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\

@@

G_{newaccount} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\

@@

The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
\begin{equation}
 \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
\end{equation}
With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
\begin{equation}
 \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{compress}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{mix}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
\end{equation}

The seed hash being:
\begin{equation}
 \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{revert}(\mathbf{n}))
\end{equation}
$E_{revert}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
\begin{equation}
 E_{revert}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
\end{equation}
We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.

>>>>

\subsection{Execution}
We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_\mathbf{i}, T_\mathbf{d}} \begin{cases} G_{\mathrm{txdatazero}} & \text{if} \quad i = 0 \\ G_{\mathrm{txdatanonzero}} & \text{otherwise} \end{cases} \\
{} & + \begin{cases} G_{\mathrm{txcreate}} & \text{if} \quad T_\mathbf{t} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
{} & + G_{\mathrm{transaction}}
\end{align}

where $T_\mathbf{i},T_\mathbf{d}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call. $G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code. $G$ is fully defined in Appendix \ref{app:fees}.

@@

If the initialization code completes successfully, a final contract-creation cost is paid, the code-deposit cost, $c$, proportional to the size of the created contract's code:
\begin{equation}
c \equiv G_{\mathrm{codedeposit}} \times |\mathbf{o}|
\end{equation}

@@

\subsection{Subtleties}
Note that while the initialisation code is executing, the newly created address exists but with no intrinsic body code. Thus any message call received by it during this time causes no code to be executed. If the initialisation execution ends with a \hyperlink{selfdestruct}{\small SELFDESTRUCT} instruction, the matter is moot since the account will be deleted before the transaction is completed. For a normal {\small STOP} code, or if the code returned is otherwise empty, then the state is left with a zombie account, and any remaining balance will be locked into the account forever.

@@

The normal halting function $H$ is defined:
\begin{equation} \label{eq:Normal Halting}
H(\boldsymbol{\mu}, I) \equiv \begin{cases}
H_{\text{\tiny RETURN}}(\boldsymbol{\mu}) & \text{if} \quad w = \text{\small RETURN} \\
() & \text{if} \quad w \in \text{\small STOP}, \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
\varnothing & \text{otherwise}
\end{cases}
\end{equation}

@@

\item[Ethereum Browser] (aka Ethereum Reference Client) A cross-platform GUI of an interface similar to a simplified browser (a la Chrome) that is able to host sandboxed applications whose backend is purely on the Ethereum protocol, which is known as Mist since 8 July 2016 (\cite{Mist}).

@@

\begin{tabularx}{\textwidth}{lrX}
Name & Value & Description* \\
$G_{\mathrm{zero}}$ & 0 & Nothing is paid for operations of the set {\small $W_{\mathrm{zero}}$}. \\
$G_{\mathrm{base}}$ & 2 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{base}}$}. \\
$G_{\mathrm{verylow}}$ & 3 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{verylow}}$}. \\
$G_{\mathrm{low}}$ & 5 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{low}}$}. \\
$G_{\mathrm{mid}}$ & 8 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{mid}}$}. \\
$G_{\mathrm{\mathrm{high}}}$ & 10 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{\mathrm{high}}}$}. \\
$G_{\mathrm{extcode}}$ & 700 & This is the amount of gas to pay for operations of the set {\small $W_{\mathrm{extcode}}$}. \\
$G_{\mathrm{balance}}$ & 400 & This is the amount of gas to pay for a {\small BALANCE} operation. \\
$G_{\mathrm{sload}}$ & 200 & This is paid for an {\small SLOAD} operation. \\
$G_{\mathrm{jumpdest}}$ & 1 & This is paid for a {\small JUMPDEST} operation. \\
$G_{\mathrm{sset}}$ & 20000 & This is paid for an {\small SSTORE} operation when the storage value is set to non-zero from zero. \\
$G_{\mathrm{sreset}}$ & 5000 & This is the amount for an {\small SSTORE} operation when the storage value's zeroness remains unchanged or is set to zero. \\
$R_{\mathrm{sclear}}$ & 15000 & This is the refund given (added into the refund counter) when the storage value is set to zero from non-zero. \hypertarget{Rselfdestruct}{}\\
$R_{\mathrm{selfdestruct}}$ & 24000 & This is the refund given (added into the refund counter) for self-destructing an account. \hypertarget{Gselfdestruct}{}\\
$G_{\mathrm{selfdestruct}}$ & 5000 & This is the amount of gas to pay for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation. \\
$G_{\mathrm{create}}$ & 32000 & This is paid for a {\small CREATE} operation. \\
$G_{\mathrm{codedeposit}}$ & 200 & This is paid per byte for a {\small CREATE} operation to succeed in placing code into the state. \\
$G_{\mathrm{call}}$ & 700 & This is paid for a {\small CALL} operation. \\
$G_{\mathrm{callvalue}}$ & 9000 & This is paid for a non-zero value transfer as part of the {\small CALL} operation. \\
$G_{\mathrm{callstipend}}$ & 2300 & This is a stipend for the called contract subtracted from $G_{\mathrm{callvalue}}$ for a non-zero value transfer. \hypertarget{Gnewaccount}{}\\
$G_{\mathrm{newaccount}}$ & 25000 & This is paid for a {\small CALL} or for a \hyperlink{selfdestruct}{\small SELFDESTRUCT} operation which creates an account. \\
$G_{\mathrm{exp}}$ & 10 & This is a partial payment for an {\small EXP} operation. \\
$G_{\mathrm{expbyte}}$ & 50 & This is a partial payment when multiplied by $\lceil\log_{256}(exponent)\rceil$ for the {\small EXP} operation. \\
$G_{\mathrm{memory}}$ & 3 & This is paid for every additional word when expanding memory. \\
$G_{\mathrm{txcreate}}$ & 32000 & This is paid by all contract-creating transactions after the {\it Homestead transition}.\\
$G_{\mathrm{txdatazero}}$ & 4 & This is paid for every zero byte of data or code for a transaction. \\
$G_{\mathrm{txdatanonzero}}$ & 68 & This is paid for every non-zero byte of data or code for a transaction. \\
$G_{\mathrm{transaction}}$ & 21000 & This is paid for every transaction. \\
$G_{\mathrm{log}}$ & 375 & This is a partial payment for a {\small LOG} operation. \\
$G_{\mathrm{logdata}}$ & 8 & This is paid for each byte in a {\small LOG} operation's data. \\
$G_{\mathrm{logtopic}}$ & 375 & This is paid for each topic of a {\small LOG} operation. \\
$G_{\mathrm{sha3}}$ & 30 & This is paid for each {\small SHA3} operation. \\
$G_{\mathrm{sha3word}}$ & 6 & This is paid for each word (rounded up) for input data to a {\small SHA3} operation. \\
$G_{\mathrm{copy}}$ & 3 & This is a partial payment for {\small *COPY} operations, multiplied by the number of words copied, rounded up. \\
$G_{\mathrm{blockhash}}$ & 20 & This is a payment for a {\small BLOCKHASH} operation. \\

@@

\begin{equation}
C(\boldsymbol{\sigma}, \boldsymbol{\mu}, I) \equiv C_{mem}(\boldsymbol{\mu}'_\mathbf{i})-C_{mem}(\boldsymbol{\mu}_\mathbf{i}) + \begin{cases}
C_\text{\tiny SSTORE}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SSTORE} \\
G_{\mathrm{exp}} & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] = 0 \\
G_{\mathrm{exp}} + G_{\mathrm{expbyte}}\times(1+\lfloor\log_{256}(\boldsymbol{\mu}_\mathbf{s}[1])\rfloor) & \text{if} \quad w = \text{\small EXP} \wedge \boldsymbol{\mu}_\mathbf{s}[1] > 0 \\
G_{\mathrm{verylow}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[2] \div 32\rceil & \text{if} \quad w = \text{\small CALLDATACOPY} \lor \text{\small CODECOPY} \\
G_{\mathrm{extcode}} + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_\mathbf{s}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_\mathbf{s}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small CALL} \lor \text{\small CALLCODE} \lor \text{\small DELEGATECALL} \\
\hyperlink{C tiny SELFDESTRUCT}{C_\text{\tiny SELFDESTRUCT}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\hyperlink{selfdestruct}{\small SELFDESTRUCT}} \\
G_{\mathrm{create}} & \text{if} \quad w = \text{\small CREATE}\\
G_{\mathrm{sha3}}+G_{\mathrm{sha3word}} \lceil \mathbf{s}[1] \div 32 \rceil & \text{if} \quad w = \text{\small SHA3}\\
G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
G_{\mathrm{sload}} & \text{if} \quad w = \text{\small SLOAD}\\
G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
G_{\mathrm{\mathrm{high}}} & \text{if} \quad w \in W_{\mathrm{\mathrm{high}}}\\
G_{\mathrm{extcode}} & \text{if} \quad w \in W_{\mathrm{extcode}}\\
G_{\mathrm{balance}} & \text{if} \quad w = \text{\small BALANCE}\\
G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small BLOCKHASH}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_\mathbf{b}[\boldsymbol{\mu}_{pc}] & \text{if} \quad \boldsymbol{\mu}_{pc} < \lVert I_\mathbf{b} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

where:
\begin{equation}
C_{mem}(a) \equiv G_{\mathrm{memory}} \cdot a + \Big\lfloor \dfrac{a^2}{512} \Big\rfloor
\end{equation}

@@

$W_{\mathrm{zero}}$ = \{{\small STOP}, {\small RETURN}\}

$W_{\mathrm{base}}$ = \{{\small ADDRESS}, {\small ORIGIN}, {\small CALLER}, {\small CALLVALUE}, {\small CALLDATASIZE}, {\small CODESIZE}, {\small GASPRICE}, {\small COINBASE},\newline \noindent\hspace*{1cm} {\small TIMESTAMP}, {\small NUMBER}, {\small DIFFICULTY}, {\small GASLIMIT}, {\small POP}, {\small PC}, {\small MSIZE}, {\small GAS}\}

$W_{\mathrm{verylow}}$ = \{{\small ADD}, {\small SUB}, {\small NOT}, {\small LT}, {\small GT}, {\small SLT}, {\small SGT}, {\small EQ}, {\small ISZERO}, {\small AND}, {\small OR}, {\small XOR}, {\small BYTE}, {\small CALLDATALOAD}, \newline \noindent\hspace*{1cm} {\small MLOAD}, {\small MSTORE}, {\small MSTORE8}, {\small PUSH*}, {\small DUP*}, {\small SWAP*}\}

$W_{\mathrm{low}}$ = \{{\small MUL}, {\small DIV}, {\small SDIV}, {\small MOD}, {\small SMOD}, {\small SIGNEXTEND}\}

$W_{\mathrm{mid}}$ = \{{\small ADDMOD}, {\small MULMOD}, {\small JUMP}\}

$W_{\mathrm{\mathrm{high}}}$ = \{{\small JUMPI}\}

$W_{\mathrm{extcode}}$ = \{{\small EXTCODESIZE}\}

Note the memory cost component, given as the product of $G_{\mathrm{memory}}$ and the maximum of 0 \& the ceiling of the number of words in size that the memory must be over the current number of words, $\boldsymbol{\mu}_\mathbf{i}$ in order that all accesses reference valid memory whether for read or write. Such accesses must be for non-zero number of bytes.

@@

G_{\mathrm{sset}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] \neq 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] = 0 \\
G_{\mathrm{sreset}} & \text{otherwise}

@@

R_{\mathrm{sclear}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[1] = 0 \; \wedge \; \boldsymbol{\sigma}[I_\mathbf{a}]_\mathbf{s}[\boldsymbol{\mu}_\mathbf{s}[0]] \neq 0 \\

@@

C_{\text{\tiny GASCAP}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) + G_{\mathrm{callstipend}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\

@@

&&&& $C_{\text{\tiny EXTRA}}(\boldsymbol{\sigma}, \boldsymbol{\mu}) \equiv G_{\mathrm{call}} + C_{\text{\tiny XFER}}(\boldsymbol{\mu}) + C_{\text{\tiny NEW}}(\boldsymbol{\sigma}, \boldsymbol{\mu})$\\

@@

G_{\mathrm{callvalue}} & \text{if} \quad \boldsymbol{\mu}_\mathbf{s}[2] \neq 0 \\

@@

G_{\mathrm{newaccount}} & \text{if} \quad \boldsymbol{\sigma}[\boldsymbol{\mu}_\mathbf{s}[1] \mod 2^{160}] = \varnothing \\

@@

The PoW-function returns an array with the compressed mix as its first item and the Keccak-256 hash of the concatenation of the compressed mix with the seed hash as the second item:
\begin{equation}
 \mathtt{PoW}(H_{\hyperlink{h cancel n}{\hcancel{n}}}, \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}) = \lbrace \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d}), \texttt{KEC}(\mathbf{s}_\mathbf{h}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}) + \mathbf{m}_\mathbf{c}(\mathtt{\small KEC}(\mathtt{\small RLP}(L_H(H_{\hcancel{n}}))), \hyperlink{H n}{H_\mathbf{n}}, \mathbf{d})) \rbrace
\end{equation}
With $H_{\hyperlink{h cancel n}{\hcancel{n}}}$ being the hash of the header without the \hyperlink{block nonce}{nonce}. The compressed mix $\mathbf{m}_\mathbf{c}$ is obtained as follows:
\begin{equation}
 \mathbf{m}_\mathbf{c}(\mathbf{h}, \mathbf{n}, \mathbf{d}) = E_{\mathrm{compress}}(E_{\mathrm{accesses}}(\mathbf{d}, \sum_{i = 0}^{n_{\mathrm{mix}}} \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}), -1), -4)
\end{equation}

The seed hash being:
\begin{equation}
 \mathbf{s}_\mathbf{h}(\mathbf{h}, \mathbf{n}) = \texttt{KEC512}(\mathbf{h} + E_{\mathrm{revert}}(\mathbf{n}))
\end{equation}
$E_{\mathrm{revert}}(\mathbf{n})$ returns the reverted bytes sequence of the \hyperlink{block nonce}{nonce} $\mathbf{n}$:
\begin{equation}
 E_{\mathrm{revert}}(\mathbf{n})[i] = \mathbf{n}[\lVert \mathbf{n} \rVert -i]
\end{equation}
We note that the ``$+$''-operator between two byte sequences results in the concatenation of both sequences.
